Hydras \& Co.

### 2022-05-18 doc/epsilon0-chapter.tex
revision book (chapter 4)

\vspace{4pt}
\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#succ}{Epsilon0.T1}}

@@

\begin{exercise}[\gaiasign]
  Look for the \gaia-theorem which corresponds to \texttt{T1is\_succ\_iff}.
\end{exercise}

@@

\begin{remark}
\label{remark:a3pat}
The alternate proof of well-foundedness using \'Evelyne Contejean's work on
    recursive path ordering~\cite{DershowitzRPO, a3pat} is available in the
    library \href{../theories/html/hydras.Epsilon0.Epsilon0rpo.html}{Epsilon0.Epsilon0rpo}.
 \end{remark}

@@

We build an instance of \texttt{ON}, and prove its correctness w.r.t. Schutte's model (see Chapter~\ref{chap:schutte}).

@@

\emph{From Module~\href{../theories/html/hydras.Schutte.Correctness_E0.html}{Schutte.Correctness\_E0}}

@@

    \section{A refinement of \texttt{E0} : an ordinal notation for \texorpdfstring{$\omega^\omega$}{omega\^omega}}

@@

\begin{exercise}[not trivial]
  It may be interesting to write a \emph{direct} proof  of well-foundedness of the order in $\omega^\omega$ (\emph{i.e.}
  without using properties of $\epsilon_0$). This exercise may
  help to understand better the proof structure of Sect.~\vref{sec:strongly-accessible}.
    \end{exercise}
%ici
    \section{A variant for hydra battles}

@@

    We prove  the commutativity of $\oplus$ in two steps. 
    First, we prove by transfinite induction on $\alpha$ that the restriction of $\oplus$ to the
    interval $[0..\alpha)$ is commutative.

@@

    So, a  natural question is `` Does there exist any strictly decreasing variant mapping
    type \texttt{Hydra} into some interval $[0,\alpha)$ (where $\alpha <\epsilon_0$) for proving the termination of all hydra battles''. The next chapter is dedicated to a formal proof that there exists no such $\alpha$, even if we consider a restriction to the set of ``standard'' battles.

>>>>

\vspace{4pt}
\noindent
\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#succ}{Epsilon0.T1}}

@@

\begin{exercise}
  \gaiasign Look for the \gaia-theorem which corresponds to \texttt{T1is\_succ\_iff}.
\end{exercise}

@@

\begin{remark}[Related work]
\label{remark:a3pat}
A proof of well-foundedness using \'Evelyne Contejean's work on
    recursive path ordering~\cite{DershowitzRPO, a3pat} is also available in the
    library \href{../theories/html/hydras.Epsilon0.Epsilon0rpo.html}{Epsilon0.Epsilon0rpo}.

    In~\cite{Manolios2005}, Manolios and Vroom prove the well-foundedness of ordinal terms below $\epsilon_0$ by reduction to the natural order on the set of natural numbers.
 \end{remark}

@@

We are now able to build an instance of \texttt{ON}.

\vspace{4pt}
\noindent
\emph{From Module~\href{../theories/html/hydras.Epsilon0.E0.html}{Epsilon0.E0}}

@@

We prove also that this notation is correct w.r.t. Schutte's model (see Chapter~\ref{chap:schutte}).

\vspace{4pt}
\noindent
\emph{From Module~\href{../theories/html/hydras.Schutte.Correctness_E0.html}{Schutte.Correctness\_E0}}

@@

    \section{An ordinal notation for \texorpdfstring{$\omega^\omega$}{omega\^omega}}

@@

\begin{exercise}
  It may be interesting to write a \emph{direct} proof  of well-foundedness of the order in $\omega^\omega$ (\emph{i.e.}
  without using properties of $\epsilon_0$). This exercise may
  help to understand better the proof structure of Sect.~\vref{sec:strongly-accessible}.
    \end{exercise}
%ici
    \section{A variant for hydra battles}

@@

    We prove  the commutativity of $\oplus$ in two steps. 
    First, we prove by transfinite induction on $\alpha$ that the restriction of $\oplus$ to the
    interval $[0,\alpha)$ is commutative.

@@

    So, a  natural question is `` Does there exist any strictly decreasing variant mapping
    type \texttt{Hydra} into some interval $[0,\mu)$ (where $\mu <\epsilon_0$) for proving the termination of all hydra battles''. The next chapter is dedicated to a formal proof that there exists no such $\mu$, even if we consider a restriction to the set of ``standard'' battles.
